Simulate

### 2023-05-31 Code/BatchEffect.Rmd
Updates from today's meeting

```{r}
two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = men_pooled, family = Gamma(link = "log"))
one_effect <- glmer(RxnTime ~ (1 | Year) , data = men_pooled, family = Gamma(link = "log"))
no_effect <- glm(RxnTime ~ 1, data = men_pooled, family = Gamma(link = "log"))
batch <- glmer(RxnTime ~ (1 | Batch) , data = men_pooled, family = Gamma(link = "log"))
anova(batch, two_effect)
print("------------------")
anova(one_effect, two_effect)
print("--------------")
par(mfrow=c(1,1))
qqnorm(resid(two_effect))
par(mfrow=c(2,1))
qqnorm(ranef(two_effect)$Year[,1])
qqnorm(ranef(two_effect)$Batch[,1])
qqnorm(ranef(batch)$Batch[,1])
qqnorm(resid(one_effect), main = "Without Batch Effect")
#anova(one_effect, two_effect)
```

>>>>

```{r}
two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = men_pooled, family = Gamma(link = "log"))
one_effect <- glmer(RxnTime ~ (1 | Year) , data = men_pooled, family = Gamma(link = "log"))
no_effect <- glm(RxnTime ~ 1, data = men_pooled, family = Gamma(link = "log"))
batch <- glmer(RxnTime ~ (1 | Batch) , data = men_pooled, family = Gamma(link = "log"))
anova(batch, two_effect)
print("------------------")
anova(one_effect, two_effect)
print("--------------")
par(mfrow=c(1,1))
qqnorm(resid(two_effect))
par(mfrow=c(2,1))
qqnorm(ranef(two_effect)$Year[,1])
qqnorm(ranef(two_effect)$Batch[,1])
qqnorm(ranef(batch)$Batch[,1])
qqnorm(resid(one_effect), main = "Without Batch Effect")
#anova(one_effect, two_effect)
```
```{r}
nested <- glmer(RxnTime ~ (1|Year) + (1|Batch:Year), data=men_pooled, family = Gamma(link="log"))
anova(nested, one_effect)
batch_year <- glmer(RxnTime ~ (1|Batch:Year), data=men_pooled, family = Gamma(link="log"))
anova(batch_year, nested)
ranef(nested)
```

### 2023-05-31 Code/Simulation.Rmd
Updates from today's meeting

```{}
The probability of observing a reaction time less than 0.01 is 0.0068649
The probability of observing a reaction time less than 0.095 is 0.0030386
The probability of observing a reaction time less than 0.09 is 0.00121

@@

```{r}
library(lme4)
set.seed(1)
simulation <- function(model, base, RxnTime){
    simfit <- function(model, n = 10000000) {
    vc <- as.data.frame(VarCorr(model))
    ## mu <- exp(coef(model)$Year[,]) ## using fitted random effects only
    z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
    mu <- exp(fixef(model) + z)
    shape <- 1 / vc$vcov[2]
    x <- rgamma(n, shape = shape, scale = mu / shape)
    x
    }
  sbase <- summary(base)
  #print(sbase)
  x <- simfit(model)
  print(1/mean(x < RxnTime))
  print(quantile(x, prob = c(.00001, .0001, .0005, .001)))
  ## compare with and without random effect
  plot(density(x)) # with random effect
  curve(dgamma(x, shape = 1 / sbase$dispersion, 
               scale = exp(coef(base)) * sbase$dispersion), 
        0.08, 0.22, add = TRUE, col = "blue") # fixed effect only
}
simulation(men_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.089)
```

@@

  sbase <- summary(base)
  x <- simfit(model)
  print(quantile(x, probs = prob))

  plot(density(x), main = "Density Plot", xlab = "RxnTime", ylab = "Density")
  curve(dgamma(x, shape = 1 / sbase$dispersion, scale = exp(coef(base)) * sbase$dispersion), 
        from = min(x), to = max(x), add = TRUE, col = "blue")
}

simulation(men_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.001)
```

```{}
Currently mean(x < 0.089) returns a proportion of x values that are less than 0.089
We instead want to set the proportion to be some set value, ex: 0.001
```

>>>>

```{}

@@

```{r}
library(lme4)
set.seed(1)
simulation <- function(model, base, RxnTime){
    simfit <- function(model, n = 10000000) {
    vc <- as.data.frame(VarCorr(model))
    ## mu <- exp(coef(model)$Year[,]) ## using fitted random effects only
    z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
    mu <- exp(fixef(model) + z)
    shape <- 1 / vc$vcov[2]
    x <- rgamma(n, shape = shape, scale = mu / shape)
    x
    }
  sbase <- summary(base)
  #print(sbase)
  x <- simfit(model)
  print(mean(x < RxnTime))
  print(quantile(x, prob = c(.00001, .0001, .0005, .001)))
  ## compare with and without random effect
  plot(density(x)) # with random effect
  curve(dgamma(x, shape = 1 / sbase$dispersion, 
               scale = exp(coef(base)) * sbase$dispersion), 
        0.08, 0.22, add = TRUE, col = "blue") # fixed effect only
}
simulation(men_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.089)
```

@@

  sbase <- summary(base)
  x <- simfit(model)
  print(quantile(x, probs = prob))
}

#simulation(men_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.001)
```

```{r}
set.seed(1)
simulation(old_finals_year_gamma_log, old_finals_one_gamma_log, 0.0001)
simulation(men_finals_year_gamma_log, men_finals_one_gamma_log, 0.0001)
simulation(old_pooled_year_gamma_log, old_pooled_one_gamma_log, 0.0001)
simulation(men_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.0001)
```
